Dependent Object Types

### 2012-09-06 doc/candidacy/slides.tex
notes

\begin{frame}{Limitations of SML modules}
\begin{itemize}
\item modules are not first-class; e.g. not possible to return a
  structure from an {\tt if} expression
\item type sharing is restricted to equality between type names, not
  general type expressions; in effect, transparency / opaqueness of a
  signature cannot be fully fined-tuned
\end{itemize}
\end{frame}

\begin{frame}{ML-Like: Calculus}
grounded in type theory; based on Girard's $F_\omega$, adding:
\begin{itemize}
\item {\it translucent sums} to model modules
\item dependent functions to model functors
\item a notion of subtyping to model module implementation-interface matching
\end{itemize}
\end{frame}

@@

\begin{frame}{ML-Like: Path-Dependent Types}
\begin{itemize}
\item Scala's abstract types  $\approx$ ML's abstract types of signatures; limitations:
\begin{itemize}
\item recursive references
\item bounded quantifications
\end{itemize}
\item translucent sum can be given a more precise type by referring to the name of its type member with a path selection; essential for %example?
\begin{itemize}
\item breaking the dependencies between (sub)fields
\item propagating typing information
\end{itemize}
\end{itemize}
\end{frame}

@@

@@

\begin{frame}{${\nu}Obj$: Details}
\begin{itemize}
\item type bindings
\begin{description}[abstract types]
\item[type aliases] $=$
\item[new types] $\expand$
\item[abstract types] $<:$
\end{description}
\item class template $\cls{x: S \sep \tup d}$ related to translucent sum; some differences:
\begin{itemize}
\item needs to be instantiated
\item ``contractive'' restriction
\end{itemize}
\item class type $\cls{x: S \sep \tup D}$
\begin{itemize}
\item contains as values classes that instantiate to objects of (sub)type $\obj{x \sep \tup D}$
\item explicit self type $S$ may be different from $\obj{x \sep \tup D}$
\item declarations in $S$ which are not in $\tup D$ play the role of abstract members, defined by mixin composition during instantiation
\end{itemize}
\end{itemize}
\end{frame}

>>>>

\begin{frame}{Limitations of SML modules}
\begin{itemize}
\item modules are not first-class; e.g. not possible to return a
  structure from an {\tt if} expression
\item type sharing is restricted to equality between type names, not
  general type expressions; in effect, transparency / opaqueness of a
  signature cannot be fully fined-tuned % example?
\end{itemize}
\end{frame}

\begin{frame}{ML-Like: Calculus}
grounded in type theory; based on Girard's $F_\omega$, adding: % quick summary of F_\omega
\begin{itemize}
\item {\it translucent sums} to model modules
\item dependent functions to model functors
\item a notion of subtyping to model module implementation-interface matching
\end{itemize}
\end{frame}

@@

\begin{frame}{ML-Like: Path-Dependent Types}
\begin{itemize}
\item Scala's abstract types  $\approx$ ML's abstract types of signatures; limitations:
\begin{itemize}
\item recursive references
\item bounded quantifications
\end{itemize}
\item translucent sum can be given a more precise type by referring to the name of its type member with a path selection; essential for %example with justification for two items below?
\begin{itemize}
\item breaking the dependencies between (sub)fields
\item propagating typing information
\end{itemize}
\end{itemize}
\end{frame}

@@

%maybe one slide on implementation?

@@

\begin{frame}{${\nu}Obj$: Details}
\begin{itemize}
\item type bindings
\begin{description}[abstract types] %examples?
\item[type aliases] $=$
\item[new types] $\expand$
\item[abstract types] $<:$
\end{description}
\item class template $\cls{x: S \sep \tup d}$ related to translucent sum; some differences:
\begin{itemize}
\item needs to be instantiated
\item ``contractive'' restriction
\end{itemize}
\item class type $\cls{x: S \sep \tup D}$
\begin{itemize}
\item contains as values classes that instantiate to objects of (sub)type $\obj{x \sep \tup D}$
\item explicit self type $S$ may be different from $\obj{x \sep \tup D}$
\item declarations in $S$ which are not in $\tup D$ play the role of abstract members, defined by mixin composition during instantiation
\end{itemize}
\end{itemize}
\end{frame}
